% Trust
% TPRC 2016 Submission
%
% Spring 2016
%
% Andy Sayler

\documentclass[11pt,letterpaper]{article}

% System Packages
\usepackage{epsfig}
\usepackage{float}
\usepackage[dvipsnames]{xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabu}
\usepackage{hyperref}
\usepackage{url}

% Local Packages
% None

% Package Options
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

% Macros
\input{macros.tex}

% Other Options
\clubpenalty = 10000
\widowpenalty = 10000

% Start
\begin{document}

\title{Categorizing, Analyzing, and Mitigating Failures of Third Party Trust}

\author{BLINDED FOR REVIEW}

\date{}

\maketitle

\begin{abstract}

From system administers to service providers to device manufactures,
whom computer users must trust and the capabilities with which they
must trust them lies at the heart of many computer security
discussions. The rise of the ``cloud'' as the preferred platform for
most modern computing applications makes questions of trust even more
critical and pressing. The modern computing landscape requires user to
trust a variety of third parties to complete even the most basic of
digital operations. Unfortunately, such trust often turns out to be
misplaced, leading to data leaks, questionable surveillance practices,
and a wide range of related privacy-harming events.

It is desirable from a public policy standpoint to understand the
nature of modern third party trust and to pursue policies that help
individuals control such trust and that minimize the likelihood of
such trust being violated. Toward these ends, this paper presents a
model for analyzing third party trust and the likelihood of trust
violations. It applies this model to the analysis of a variety of well
publicized data breaches that have occurred over the previous ten
years and uses it to categorize the common manners in which third
parties violate the trust user place in them. Finally, this paper
presents a number of proposed techniques, both technological and
policy-based, for helping to mitigate the degree of trust users must
place in many third parties as well as the likelihood that these
parties will violate this trust.

\end{abstract}

\input{sec_intro}
\input{sec_model}
\input{sec_analysis}
\input{sec_mitigation}
\input{sec_conclusion}

\bibliographystyle{acm}
\bibliography{refs}

\end{document}
